---
title: "Problem Set 4"
author: "Christy Smith"
date: "3/9/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#>")
```

# Submission

1. Fork this repository to your own account
1. Make changes to the `README.Rmd` file (including the author field above).
1. Knit the file to HTML format
1. Publish HTML on Rpubs (you will need to sign up for the free account) and *add the link below*.
1. Submit a PR with your answers.

*Link to Rpubs document:* http://rpubs.com/chrismit258/258789

# Overview

Take the [datacamp course on joining data](https://www.datacamp.com/courses/joining-data-in-r-with-dplyr) to learn the join verbs.

You will analyze some data in the [`nihexporter`](https://github.com/jayhesselberth/nihexporter) package to answer some questions. The questions will cover methods in `dplyr` including two-table verbs.

First you need to install the `nihexporter` package (and `devtools`, if you don't have it already).

```r
# install.packages('devtools')
devtools::install_github("jayhesselberth/nihexporter")
```

Read the [Overview vignette](http://jayhesselberth.github.io/nihexporter/articles/nihexporter.html) to learn the structure of the pacakge. Take a look at two tables from the package:

```{r overview}
library(nihexporter)
library(tidyverse)
library(broom)

projects
project_io
```

# Problems
#Notes to self: project.num is the main ID that you can use to link tables together. "RCR" refers to relative citation ratio (impact factor for a manuscript). 

## Problem 1

The `projects` and `project_io` tables have the `r intersect(names(projects), names(project_io))` column in common. Use this column to perform a `left_join` and identify the most expensive project for each institute. Project institutes are in the `projects` table, whereas total project costs are available in the `project_io` table. Note you will have to use the `arrange` and `slice` verbs from `dplyr`.

```{r problem_1}
projects %>% 
  select(project.num, institute) %>% 
  left_join(project_io) %>% 
  select(project.num, institute, total.cost) %>% 
  group_by(institute) %>% 
  arrange(desc(total.cost)) %>% 
  slice(1:1) %>% 
  arrange(desc(total.cost))
```

## Problem 2

Create a plot that describes the change in output (number of publications) for the most expensive grant from the above list. Note that you cannot use `n.pubs` for this, as that is an aggregate count over time. You need a tbl of following structure, which you can then use to calculate the number of publications each year (hint: use `n()` in a `summarize` statement: 

```{r}
projects %>%
  head() %>%
  left_join(publinks) %>%
  left_join(publications) %>%
  select(project.num, fiscal.year, fy.cost, pmid) %>%
  na.omit()
```


```{r problem_2}
projects %>% 
  select(project.num, institute) %>% 
  left_join(project_io) %>% 
  select(project.num, institute, total.cost) %>% 
  group_by(institute) %>% 
  arrange(desc(total.cost)) %>% 
  slice(1:1) %>% 
  arrange(desc(total.cost)) %>% 
  ungroup %>% 
  slice(1:1) -> most_expens_proj

projects %>% 
  semi_join(most_expens_proj) %>% 
  left_join(publinks) %>% 
  left_join(publications) %>% 
  select(project.num, pub.year, fy.cost, pmid) %>% 
  group_by(pub.year) %>% 
  summarize(total.pubs =n()) %>% 
  ggplot(aes(x=pub.year, y=total.pubs)) + geom_point()
```

## Problem 3

Use the `projects` and `clinical_studies` table to calculate how many clinical studies have been started by projects from each institute.

```{r problem_3}
projects %>%
  select(institute, project.num) %>%
  left_join(clinical_studies) %>%
  select(institute, project.num, trial.id) %>%
  filter(!is.na(trial.id)) %>%
  group_by(institute) %>%
  summarise(n_distinct(trial.id))
```

## Problem 4

Calculate correlations between `total.cost` and `n.pubs` for each institute. You will need to `join` the `projects` and `project_io` tables to link `institute` with `total.cost` and `n.pubs`. You will then need to use `group_by` and `do` from `dplyr` and `tidy` from `broom` to calculate correlations.

```{r problem_4}
projects %>%
  left_join(project_io) %>%
  select(institute, n.pubs, total.cost) %>%
  filter(!is.na(n.pubs), !is.na(total.cost)) %>%
  group_by(institute) %>%
  do(tidy(cor.test(.$n.pubs, .$total.cost)))
```

## Problem 5 (extra credit)

Correlate [`rcr`](https://icite.od.nih.gov/) values from the `publications` table to one of:

- total cost (overall for NIH, or by institute)
- budget mechanism (e.g., RFA, PAR, etc.)
- type of grant (e.g. R01, P01, R21, etc.). 

#I used total cost per institute. 

```{r extra_credit}
projects %>%
  left_join(project_io) %>%
  select(institute, project.num, total.cost) %>%
  left_join(publinks) %>%
  filter(!is.na(pmid)) %>%
  left_join(publications) %>%
  select(institute, total.cost, rcr) %>%
  filter(!is.na(rcr), !is.na(total.cost)) %>%
  group_by(institute) %>%
  do(tidy(cor.test(.$total.cost, .$rcr)))

```

## Problem 6 (extra extra credit)

Use [`gganimate`](https://github.com/dgrtwo/gganimate) to illustrate the change in cost of R01s for a few institutes over time.


```{r extra_extra_credit}
library(gapminder)
library(ggplot2)
library(gganimate)

projects %>%
  filter(activity == "R01") %>%
  select(institute, project.num, fiscal.year) %>%
  left_join(project_io) %>%
  filter(institute == "AI" | institute == "CA" | institute == "HD") %>%
  select(institute, fiscal.year, total.cost) %>%
  filter(!is.na(total.cost)) %>%
  group_by(institute, fiscal.year) %>%
  summarize(cost = sum(total.cost)) %>%
  ggplot(aes(x=fiscal.year, y=cost, color=institute, frame=fiscal.year)) + geom_point() -> p

gganimate(p)

```
